---
title: "Week 3 assignment"
author : "Jihan Yu"
Date : "April 28, 2020"
output: ioslides_presentation
---
## Scatter plots of diamonds data
### April, 28, 2020

  This document shows scatter plots of the diamonds data in the ggplot2 library. The axes are designated as follow: X-axis is price, and Y-axis is carat. The color of the data is assigned as cut, color or clarity.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, fig.align = "center", message=FALSE, warning=FALSE, cache=TRUE)
```

```{r data pre-proess}
library(plotly); 
data(diamonds)
```

## 3D scatter plot(1)
### Price vs Carat vs cut (color : cut)
```{r price vs carat vs cut}
plot_ly(diamonds, x = ~price, y = ~carat, z = ~cut, color = ~cut)
```

## 3D scatter plot(2)
### Price vs Carat vs carat (color : carat)
```{r price vs carat vs color}
plot_ly(diamonds, x = ~price, y = ~carat, z = ~color, color = ~color)
```

## 3D scatter plot(3)
### Price vs Carat vs clarity (color : clarity)
```{r price vs carat vs clarity}
plot_ly(diamonds, x = ~price, y = ~carat, z = ~clarity, color = ~clarity)
```